### 2020-07-04 msr.tex
final edits for 2nd draft

The first \gls{MSR}, named the \gls{ARE}, dates back to the 1940s
as part of the US Aircraft Nuclear Propulsion program
\cite{rosenthal_molten-salt_1970}. Researchers recommended the molten fluoride
salts in particular for high uranium solubility, chemical stability, low vapor
pressure even at high temperatures, decent heat transfer properties,
resistance against radiation damage, and reduced corrosive effects on some
common structural material \cite{rosenthal_molten-salt_1970}. They
subsequently built the 2.5 MW$_{\text{th}}$ ARE reactor at \gls{ORNL}, where
it achieved criticality on November 1954 and generated 100 MWh over nine days.
The fuel consisted of enriched uranium in a molten salt mixture of NaF,
ZrF$_4$, and UF$_4$, and the reactor used blocks of BeO for neutron
moderation. The aircraft program
ultimately never came to fruition as the development of intercontinental
ballistic missiles effectively eliminated the need for long-range
nuclear-powered bomber aircraft.

@@

Building on their experience with the \gls{MSRE}, \gls{ORNL} proposed a
new program for the construction and operation of a demonstration reactor
based on the Molten Salt Breeder Reactor (MSBR) concept that they had
developed \cite{macpherson_molten_1985}. The \gls{MSBR} is a thermal-spectrum,
single fluid reactor with fertile $^{232}$Th isotopes mixed directly into the
FLiBe molten salt for $^{233}$U breeding \cite{gehin_liquid_2016}. Like the
\gls{MSRE}, the \gls{MSBR} relies on continuous online reprocessing to add
fertile material and remove fission product neutron poisons. Researchers
estimated the doubling time (the minimum amount of time required to produce
enough fissile material to start up another \gls{MSBR}) to be
approximately 22 years. However, \gls{ORNL} failed to secure funding for the
new program in their two attempts in 1972 and 1974. Nevertheless, from a
technical perspective, two independent
technology evaluation and design studies of the \gls{MSR} had reported
favorably on the promise of the system \cite{macpherson_molten_1985}.

In spite of this setback, research into \glspl{MSR} continued through the late
1970s. In 1980, \gls{ORNL} published a report describing a new \gls{MSR}
concept, called the \gls{DMSR} \cite{gehin_liquid_2016} with denatured
$^{235}$U fuel (i.e. low-enriched uranium). The \gls{ORNL} researchers
developed this design in response to the fuel reprocessing restrictions
introduced by President Ford in 1976. The \gls{DMSR} would operate as a
once-through
converter system without fuel reprocessing. While the fuel largely consists of
19.75 \% \gls{HALEU}, the initial core loading includes thorium to
boost its conversion ratio throughout its lifetime. It has a continuous
online feed consisting of \gls{HALEU} to maintain criticality, and denatured
$^{235}$U to keep uranium enrichment levels below nuclear non-proliferation
policy thresholds. The design also includes a gas sparging system for removing
gaseous fission products, while noble metals plate out onto the walls of
the coolant loop. The older \gls{MSBR} design had a significant drawback; the
extensive neutron damage in the graphite moderator necessitated frequent
replacement (every four years) throughout its operational lifetime. The
\gls{DMSR} avoids this issue by having a lower power density while maintaining
the overall power output of 2250 MW$_{\text{th}}$. As a result, researchers
projected that the graphite moderator would last for the entirety of the
\gls{DMSR}'s design lifetime.

There was a concurrent program at the UK Atomic Energy Authority for the
development of a 2500 MW$_{\text{e}}$ lead-cooled
\gls{MCFR} concept \cite{smith_assessment_1974}. It is a dual fluid system,
with separate loops for the fuel salt and the blanket salt. The blanket is a
1 m-wide tank surrounding the core. The absence of moderators and the choice
of chloride over fluoride salt resulted in a relatively hard neutron spectrum
which favors $^{239}$Pu breeding over the thorium cycle. The UK researchers
performed some experiments to study molten salt chemistry but there were no
reactor prototypes. The UK program eventually shut down just like its
US counterpart partly due to the successful demonstration of the Prototype
Fast Reactor which had achieved criticality in 1974.

Following a lull lasting through the late 20th century, \gls{MSR} research
picked up pace due to renewed interest initiated by the Generation IV Forum in
2001. Today, there are numerous \gls{MSR} concepts under active development
led by various national and commercial bodies.

@@

\glspl{MSR} rely on natural physical phenomena for passive safety such as the
strong negative temperature reactivity feedback of the fuel salt due to
greater temperature-induced expansion in liquid fuel than solid
fuel. Combined with the Doppler broadening of resonance capture cross sections
present in both fuel forms, there would be a smaller temperature
increase following an unprotected reactivity insertion. The overall
temperature reactivity coefficient varies widely among \gls{MSR}
designs due to other structures, such as moderators and reflectors, present in
the core. In particular, graphite moderators tend to have slightly positive
temperature reactivity coefficients. \gls{MSBR} concept has this issue, but
the total temperature reactivity coefficient is still relatively large and
negative \cite{rykhlevskii_modeling_2019}. The negative temperature reactivity
feedback provides a large degree of control and stability as it is always
present in an \gls{MSR} regardless of the operating conditions.

Continuous online fuel reprocessing allows operators to maintain low excess
reactivity inventories in the core as additional fuel can be added on an ad
hoc basis. Reprocessing and gas sparging systems help reduce fissile
requirements by continuously removing neutron poisons. These factors, in
addition to the strong negative temperature
reactivity coefficient, diminish the likelihood and severity of unprotected
criticality accidents in \glspl{MSR} \cite{elsheikh_safety_2013}. In the
unlikely situation in which an \gls{MSR} encounters a severe runaway reaction,
\glspl{MSR} rely on another passive safety feature: freeze plugs. Various
freeze plugs designs exist for different \glspl{MSR}. The freeze plug
in the \gls{MSFR} concept is a plug of solidified salt at the bottom of the
core actively cooled by fans or other cooling systems to keep its temperature
just below the freezing point of the salt \cite{aji_experimental_2020}. In
general, when temperatures in the core exceed a certain threshold during a
dangerous transient, the freeze plug melts and the molten salt in the core
drains into a containment tank designed to keep the salt in a subcritical
configuration. This is especially easy to achieve in thermal-spectrum
\glspl{MSR} as the absence of moderators in the containment tank would
automatically drive the multiplication factor down below unity
\cite{elsheikh_safety_2013}.

@@

\glspl{MSR} also typically have large margin-to-boiling under nominal
operating conditions so that fuel salt boiling does not occur
\cite{elsheikh_safety_2013}. Furthermore, the reactor vessel is consequently
subject to much lower stresses as \glspl{MSR} operate at near-atmospheric
pressure levels. Thus, the probability of pipe ruptures due to high pressure
is low.

However, there are also some safety disadvantages associated with \glspl{MSR}.
Firstly, \glspl{MSR} have smaller fractions of delayed neutron precursors in
the active core region as some of them decay in the external loop regions.
This complicates reactor control and may result in faster transients due to
the decrease in average neutron lifetime. Secondly, hot molten salts are
generally corrosive and the corrosion mechanism is different from conventional
corrosion induced by water and other common agents \cite{yoshioka_7_2017}. The
intense gamma and neutron radiation in the core may also accelerate corrosion
in the structural components. Lastly, overcooling in pipes and heat exchangers
may pose operational challenges as the salt can freeze and restrict flow,
thereby causing a loss of flow accident \cite{ho_molten_2013}.

@@

\glspl{MSR} have several positive sustainability features.
\glspl{MSR} have good neutron economy due to the continuous removal of
neutron poisons by the online reprocessing and gas sparging systems
\cite{kamei_recent_2012}. Both
$^{232}$Th/$^{233}$U and $^{238}$U/$^{239}$Pu fuel cycles are viable
candidates for breeding in \glspl{MSR}, with the former being more suited for
thermal reactors and the latter being more suited for fast reactors.

The $^{232}$Th/$^{233}$U fuel cycle produces significantly less \gls{TRU}
waste than the other cycles due to the smaller atomic masses of $^{232}$Th and
$^{233}$U. This reduces the overall radiotoxicity and long-term decay heat
associated with long-lived plutonium and \gls{MA} isotopes. The combination
of \gls{TRU} fuel and $^{232}$Th feed in fast spectrum \glspl{MSR} would help
reduce overall levels of \gls{TRU} waste going into long-term storage in
nuclear waste repositories \cite{merle-lucotte_launching_2011}.

Nuclear non-proliferation concerns in thorium-fueled \glspl{MSR} involve the
separation of the intermediate $^{233}$Pa
isotope from the fuel salt. $^{233}$Pa decays into $^{233}$U with a half-life
of approximately 27 days and the $^{233}$U produced is equivalent in potency
to $^{239}$Pu for nuclear weapons production \cite{grape_10_2017}. The highly
radioactive $^{232}$U
by-product provides some level of proliferation resistance but nuclear
proliferators can sidestep this complication by
separating the $^{232}$U away from the combined $^{232}$Pa/$^{233}$Pa stream
after most of the $^{232}$Pa has decayed into $^{232}$U at a short half-life
of 1.31 days \cite{grape_10_2017}. Safeguards by design and close monitoring
of \glspl{MSR} are therefore essential to avoiding spent nuclear fuel
diversion.

Economic analyses of \glspl{MSR} are preliminary at the current
stage of \gls{MSR} development. Qualitatively, many technical factors favor
\glspl{MSR} over \gls{LWR}. Some of these factors include: smaller reactor
core size due to low operating pressures, higher thermal efficiency and
cheaper air-cooling systems due to high operating temperatures, reduction in
fuel fabrication costs, and shorter reactor shutdown periods due to online
refueling \cite{dolan_1_2017}. Moir \cite{moir_recommendations_2008}
published data from \gls{ORNL} reports on the estimated electricity costs for
a 1000 MWe \gls{MSR} with 20\% enriched $^{235}$U fuel and no reprocessing
throughout the reactor's 30-year lifetime. Adjusted to year 2000 dollars, the
electricity costs for this type of \gls{MSR} was 7\% and 8\% lower than that
of pressurized water reactor and coal power plants. These differences are
within the $\pm$10\% uncertainty range but there is room for some optimism.
The author also notes that costs could be lower in an \gls{MSR} with
fertile feed and online reprocessing could reach fuel self sufficiency.

In terms of technological readiness, \glspl{MSR} still require significant
R\&D efforts and engineering demonstrations for experimental validation of
various components before a full commercial model can be commissioned. Work
towards creating a safety and licensing framework for \glspl{MSR} has picked
up pace only in recent years due to the growing interest from commercial
\gls{MSR} developers.

@@

In the past two decades, researchers have developed several
new tools for simulating steady-state and transient behavior in \glspl{MSR}.
Many of the earlier efforts featured simplifications in simulating
thermal-hydraulics by solving 1-D Navier-Stokes equations or using
predetermined uniform velocity fields \cite{krepel_dyn3d-msr_2007,
kophazi_development_2009}. In more recent years, there has been
significant progress towards fully coupled, spatial codes that feature
2-D axisymmetric or full 3-D models. In 2011, Cammi et al.
\cite{cammi_multi-physics_2011} performed a ``\gls{MPM}'' analysis
of a simplified 2-D axisymmetric model of a single \gls{MSBR} fuel channel
using the commercial finite element analysis software COMSOL Multiphysics. The
physics were implemented through the two-group neutron diffusion equations,
and the \gls{RANS} standard $k$-$\epsilon$ turbulence model, for the
neutronics and thermal-hydraulics respectively. The
authors emphasized the need for proper full coupling of the multiphysics and
presented both steady-state and transient results in various
scenarios such as reactivity insertions, changes in pumping rate, and the
presence of periodic perturbations. This approach was featured again in a
later paper by Fiorina et al. \cite{fiorina_modelling_2014} in 2014 for a 2-D
axisymmetric model of the
\gls{MSFR}. The authors presented results from the Politecnico di Milano
COMSOL-based approach, and another approach by researchers from Delft
University of Technology, in which they coupled their in-house neutronics and
thermal-hydraulics codes, DALTON and HEAT respectively. With multigroup
neutron diffusion and \gls{RANS} formulations on ultra fine meshes, both
models showed good agreement in the steady-state neutron flux, temperature,
and \gls{DNP} distributions, and in the power responses following various
accident transient initiations. Aufiero et al. \cite{aufiero_development_2014}
concurrently developed
a full-core 3-D model of the \gls{MSFR} on OpenFOAM, albeit with one-group
neutron diffusion to reduce computational load. With the 3-D model, the
authors could simulate the asymmetric reactor response to the failure of a
single pump in the sixteen-pump \gls{MSFR} configuration. The authors also
provided some quantitative data supporting the use of implicit coupling over
explicit coupling to obtain accurate solutions of the transient cases.
Recognizing the huge computational burden required for full 3-D simulations, 
later authors came up with innovative ways to alleviate this issue such as
selective geometrical reduced order modeling for various components of a
reactor based on the importance of the physical phenomena being simulated
\cite{zanetti_geometric_2015}, or using a novel, efficient method for
neutronics calculations \cite{laureau_transient_2017}.

@@

The \gls{MSFR} is a European reference fast-spectrum \gls{MSR} concept
\cite{mathieu_thorium_2006, merle_optimized_2007}. Table \ref{table:msfr} and
Figure \ref{fig:msfr} show the main specifications and schematic view of the
\gls{MSFR}, respectively. Developed from the \gls{MSBR} design, the
\gls{MSFR} is intended to run primarily on a closed thorium fuel cycle with
continuous online fuel reprocessing. Several reasons motivated the omission of
graphite moderators from the original \gls{MSBR} design. Graphite is
susceptible to
long-term radiation damage and replacement is likely to be necessary during
the operating lifetime of the reactor. Graphite also has a positive
temperature coefficient of reactivity; eliminating graphite from the design
ensures a greater safety margin \cite{mathieu_thorium_2006}. While negative
temperature coefficients are attainable with very thermalized spectra,
breeding ratios deteriorated significantly due to parasitic absorption in the
large volume of graphite needed for thermalization
\cite{mathieu_thorium_2006}.
%
\begin{figure}[htb!] 
	\centering
	\includegraphics[width=0.8\textwidth]{MSFR}
	\caption{Schematic view of the MSFR concept. Figure reproduced from
	Brovchenko et al. \cite{brovchenko_design-related_2013}.}
	\label{fig:msfr}
\end{figure}
%
\begin{table}[htb!]
    \small
	\caption{Main specifications of the \gls{MSFR} concept
				\cite{serp_molten_2014}.}
	\centering
	\begin{tabular}{ l r }
		\hline
		Parameter & Value \\
		\hline
		Thermal/Electric output [MW$_{\text{th}}$/MW$_{\text{e}}$] & 3000 /
		1500 
		\\
		Salt volume [m$^3$] & 18 \\
		Salt fraction in core & 0.5 \\
		Number of circulation loops & 16 \\
		Nominal flow rate [kg s$^{-1}$] & 18500  \\
		Nominal circulation time [s] & 4.0 \\
		Inlet/outlet temperature [K] & 923 / 1023 \\
		Blanket volume [m$^3$] & 7.3\\
		\hline
	\end{tabular}
	\label{table:msfr}
\end{table}

>>>>

The first \gls{MSR}, named the \gls{ARE}, dates back to the 1940s
as part of the US Aircraft Nuclear Propulsion program
\cite{rosenthal_molten-salt_1970}. Researchers recommended the molten fluoride
salts in particular for high uranium solubility, chemical stability, low vapor
pressure even at high temperatures, decent heat transfer properties,
resistance against radiation damage, and reduced corrosive effects on some
common structural material \cite{rosenthal_molten-salt_1970}. They
subsequently built the 2.5 MW$_{\text{th}}$ ARE reactor at \gls{ORNL}, where
it achieved criticality on November 1954 and generated 100 MWh over nine days.
The fuel consisted of enriched uranium in a molten salt mixture of NaF,
ZrF$_4$, and UF$_4$. Additionally, the reactor used blocks of BeO for neutron
moderation. The aircraft program
ultimately never came to fruition as the development of intercontinental
ballistic missiles effectively eliminated the need for long-range
nuclear-powered bomber aircraft.

@@

Building on their experience with the \gls{MSRE}, \gls{ORNL} proposed a
new program for the construction and operation of a demonstration reactor
based on the \gls{MSBR} concept that they had
developed \cite{macpherson_molten_1985}. The \gls{MSBR} is a thermal-spectrum,
single fluid reactor with fertile $^{232}$Th isotopes mixed directly into the
FLiBe molten salt for $^{233}$U breeding \cite{gehin_liquid_2016}. Like the
\gls{MSRE}, the \gls{MSBR} relies on continuous online reprocessing to add
fertile material and remove fission product neutron poisons. Researchers
estimated the doubling time (the minimum amount of time required to produce
enough fissile material to start up another \gls{MSBR}) to be
approximately 22 years. However, \gls{ORNL} failed to secure funding for the
new program in their two attempts in 1972 and 1974. Nevertheless, from a
technical perspective, two independent
technology evaluation and design studies of the \gls{MSR} had reported
favorably on the promise of the system \cite{macpherson_molten_1985}.

In spite of this setback, research into \glspl{MSR} continued through the late
1970s. In 1980, \gls{ORNL} published a report describing a new \gls{MSR}
concept, called the \gls{DMSR} \cite{gehin_liquid_2016} with denatured
$^{235}$U fuel (i.e. low-enriched uranium). The \gls{ORNL} researchers
developed this design in response to the fuel reprocessing restrictions
introduced by President Ford in 1976. The \gls{DMSR} would operate as a
once-through
converter system without fuel reprocessing. While the fuel consists of 19.75
\% high-assay low-enriched uranium, the initial core loading includes thorium
to boost its conversion ratio throughout its lifetime. It has a continuous
online feed consisting of \gls{HALEU} to maintain criticality, and denatured
$^{235}$U to keep uranium enrichment levels below nuclear non-proliferation
policy thresholds. The design also includes a gas sparging system for removing
gaseous fission products, while noble metals plate out onto the walls of
the coolant loop. The older \gls{MSBR} design had a significant drawback; the
extensive neutron damage in the graphite moderator necessitated frequent
replacement (every four years) throughout its operational lifetime. The
\gls{DMSR} avoids this issue by having a lower power density while maintaining
the overall power output of 2250 MW$_{\text{th}}$. As a result, researchers
projected that the graphite moderator would last for the entirety of the
\gls{DMSR}'s design lifetime.

There was a concurrent program at the UK Atomic Energy Authority for the
development of a 2500 MW$_{\text{e}}$ lead-cooled Molten Chloride Fast Reactor
concept \cite{smith_assessment_1974}. It is a dual fluid system,
with separate loops for the fuel salt and the blanket salt. The blanket is a
1 m-wide tank surrounding the core. The absence of moderators and the choice
of chloride over fluoride salt resulted in a relatively hard neutron spectrum
which favors $^{239}$Pu breeding over the thorium cycle. The UK researchers
performed some experiments to study molten salt chemistry but there were no
reactor prototypes. The UK program eventually shut down just like its
US counterpart partly due to the successful demonstration of the Prototype
Fast Reactor which had achieved criticality in 1974.

Following a lull lasting through the late 20th century, \gls{MSR} research
picked up pace due to renewed interest initiated by the Generation IV Forum in
2001. Today, various national and commercial bodies are developing numerous
\gls{MSR} concepts.

@@

\glspl{MSR} rely on natural physical phenomena for passive safety such as the
strong negative temperature reactivity feedback of the fuel salt due to
greater temperature-induced expansion in liquid fuel than solid
fuel. Combined with the Doppler broadening of resonance capture cross sections
present in both fuel forms, there would be a smaller temperature
increase following an unprotected reactivity insertion. The overall
temperature reactivity coefficient varies widely among \gls{MSR}
designs due to other structures, such as moderators and reflectors, present in
the core. In particular, graphite moderators tend to have slightly positive
temperature reactivity coefficients. The \gls{MSBR} concept has this issue,
but the total temperature reactivity coefficient is still relatively large and
negative \cite{rykhlevskii_modeling_2019}. The negative temperature reactivity
feedback provides a great degree of control and stability as it is always
present in an \gls{MSR} regardless of the operating conditions.

Continuous online fuel reprocessing allows operators to maintain low excess
reactivity inventories in the core as additional fuel can be added on an ad
hoc basis. Reprocessing and gas sparging systems help reduce fissile
requirements by continuously removing neutron poisons. These factors, in
addition to the strong negative temperature
reactivity coefficient, diminish the likelihood and severity of unprotected
criticality accidents in \glspl{MSR} \cite{elsheikh_safety_2013}. In the
unlikely situation in which an \gls{MSR} encounters a severe runaway reaction,
\glspl{MSR} rely on another passive safety feature: freeze plugs. Various
freeze plugs designs exist for different \glspl{MSR}. The freeze plug
in the \gls{MSFR} concept is a plug of solidified salt at the bottom of the
core actively cooled by fans or other cooling systems to keep its temperature
just below the freezing point of the salt \cite{aji_experimental_2020}. When
temperatures in the core exceed a certain threshold during a
dangerous transient, the freeze plug melts and drains the molten salt
into a containment tank designed to keep the salt in a subcritical
configuration. This is especially easy to achieve in thermal-spectrum
\glspl{MSR} as the absence of moderators in the containment tank would
automatically drive the multiplication factor down below unity
\cite{elsheikh_safety_2013}.

@@

\glspl{MSR} also typically have a high margin to boiling under nominal
operating conditions so that fuel salt boiling is absent
\cite{elsheikh_safety_2013}. Furthermore, the reactor vessel is consequently
subject to much lower stresses as \glspl{MSR} operate at near-atmospheric
pressure levels. Thus, the probability of pipe ruptures due to high pressure
is low.

However, \glspl{MSR} do have some associated safety risks.
Firstly, \glspl{MSR} have smaller fractions of delayed neutron precursors in
the active core region as some of them decay in the external loop regions.
This complicates reactor control and may result in faster transients due to
the decrease in average neutron lifetime. Secondly, hot molten salts are
corrosive and the corrosion mechanism is different from conventional
corrosion induced by water and other common agents \cite{yoshioka_7_2017}. The
intense gamma and neutron radiation in the core may also accelerate corrosion
in the structural components. Lastly, overcooling in pipes and heat exchangers
may pose operational challenges as the salt can freeze and restrict flow,
causing a loss of flow accident \cite{ho_molten_2013}.

@@

\glspl{MSR} have several positive sustainability features.
The continuous removal of neutron poisons by the online reprocessing and gas
sparging systems creates good neutron economy in \glspl{MSR}
\cite{kamei_recent_2012}. Both
$^{232}$Th/$^{233}$U and $^{238}$U/$^{239}$Pu fuel cycles are viable
candidates for breeding in \glspl{MSR}, with the former being more suited for
thermal reactors and the latter being more suited for fast reactors.

The $^{232}$Th/$^{233}$U fuel cycle produces significantly less \gls{TRU}
waste than the other cycles due to the smaller atomic masses of $^{232}$Th and
$^{233}$U. This reduces the overall radiotoxicity and long-term decay heat
associated with long-lived plutonium and \gls{MA} isotopes. The combination
of \gls{TRU} fuel and $^{232}$Th feed in fast spectrum \glspl{MSR} contributes
to lower levels of \gls{TRU} waste going into long-term storage in
nuclear waste repositories \cite{merle-lucotte_launching_2011}.

Nuclear non-proliferation concerns in thorium-fueled \glspl{MSR} involve the
separation of the intermediate $^{233}$Pa
isotope from the fuel salt. $^{233}$Pa decays into $^{233}$U with a half-life
of approximately 27 days and the $^{233}$U produced is equivalent in potency
to $^{239}$Pu for nuclear weapons production \cite{grape_10_2017}. The highly
radioactive $^{232}$U
by-product provides some level of proliferation resistance but nuclear
proliferators can sidestep this complication; they can
separate the $^{232}$U away from the combined $^{232}$Pa/$^{233}$Pa stream
after most of the $^{232}$Pa has decayed into $^{232}$U at a short half-life
of 1.31 days \cite{grape_10_2017}. Safeguards by design and close monitoring
of \glspl{MSR} are therefore essential to avoiding spent nuclear fuel
diversion.

Economic analyses of \glspl{MSR} are preliminary at the current
stage of \gls{MSR} development. Qualitatively, many technical factors favor
\glspl{MSR} over \glspl{LWR}. Some of these factors include: smaller reactor
core size due to low operating pressures, higher thermal efficiency and
cheaper air-cooling systems due to high operating temperatures, reduction in
fuel fabrication costs, and shorter reactor shutdown periods due to online
refueling \cite{dolan_1_2017}. Moir \cite{moir_recommendations_2008}
published data from \gls{ORNL} reports on the estimated electricity costs for
a 1000 MWe \gls{MSR} with 20\% enriched $^{235}$U fuel and no reprocessing
throughout the reactor's 30-year lifetime. Adjusted to year 2000 dollars, the
electricity costs for this type of \gls{MSR} was 7\% and 8\% lower than that
of pressurized water reactors and coal power plants. These differences are
within the $\pm$10\% uncertainty range but Moir remains optimistic.
The author also notes that costs could be lower in an \gls{MSR} with
fertile feed and online reprocessing could reach fuel self sufficiency.

In terms of technological readiness, \glspl{MSR} still require significant
research efforts and engineering demonstrations for experimental validation of
various components before a full commercial model can be commissioned. Work
towards creating a safety and licensing framework for \glspl{MSR} has picked
up pace only in recent years due to the growing interest from commercial
\gls{MSR} developers.

@@

In the past two decades, researchers have developed several
new tools for simulating steady-state and transient behavior in \glspl{MSR}.
Many of the earlier efforts featured simplifications in simulating
thermal-hydraulics by solving 1-D Navier-Stokes equations or using
predetermined uniform velocity fields \cite{krepel_dyn3d-msr_2007,
kophazi_development_2009}. In more recent years, there has been
significant progress towards fully coupled, spatial codes that feature
2-D axisymmetric or full 3-D models. In 2011, Cammi et al.
\cite{cammi_multi-physics_2011} performed a ``\gls{MPM}'' analysis
of a simplified 2-D axisymmetric model of a single \gls{MSBR} fuel channel
using the commercial finite element analysis software COMSOL Multiphysics. The
physics were implemented through the two-group neutron diffusion equations,
and the \gls{RANS} standard $k$-$\epsilon$ turbulence model, for the
neutronics and thermal-hydraulics respectively. The
authors emphasized the need for proper full coupling of the multiphysics and
presented both steady-state and transient results in various
scenarios such as reactivity insertions, changes in pumping rate, and the
presence of periodic perturbations. This approach was featured again in a
later paper by Fiorina et al. \cite{fiorina_modelling_2014} in 2014 for a 2-D
axisymmetric model of the
\gls{MSFR}. The authors presented results from the Politecnico di Milano
COMSOL-based approach, and another approach by researchers from Delft
University of Technology, in which they coupled their in-house neutronics and
thermal-hydraulics codes, DALTON and HEAT respectively. With multigroup
neutron diffusion and \gls{RANS} formulations on ultra fine meshes, both
models agreed on the steady-state neutron flux, temperature,
and \gls{DNP} distributions, as well as the power responses following various
accident transient initiations. Aufiero et al. \cite{aufiero_development_2014}
concurrently developed
a full-core 3-D model of the \gls{MSFR} on OpenFOAM, albeit with one-group
neutron diffusion to reduce computational load. With the 3-D model, the
authors could simulate the asymmetric reactor response to the failure of a
single pump in the sixteen-pump \gls{MSFR} configuration. The authors also
provided quantitative data supporting the use of implicit coupling over
explicit coupling to obtain accurate solutions of the transient cases.
Recognizing the huge computational burden required for full 3-D simulations, 
later authors came up with innovative ways to alleviate this issue. Methods
include
selective geometrical reduced order modeling for various components of a
reactor based on the importance of the physical phenomena being simulated
\cite{zanetti_geometric_2015}, or using a novel, efficient method for
neutronics calculations \cite{laureau_transient_2017}.

@@

The \gls{MSFR} is a European reference fast-spectrum \gls{MSR} concept
\cite{mathieu_thorium_2006, merle_optimized_2007}. Table \ref{table:msfr} and
Figure \ref{fig:msfr} show the main specifications and schematic view of the
\gls{MSFR}, respectively. Developed from the \gls{MSBR} design, the
\gls{MSFR} is intended to run primarily on a closed thorium fuel cycle with
continuous online fuel reprocessing. Several reasons motivated the omission of
graphite moderators from the original \gls{MSBR} design. Graphite is
susceptible to
long-term radiation damage and replacement is likely to be necessary during
the operating lifetime of the reactor. Graphite also has a positive
temperature coefficient of reactivity; eliminating graphite from the design
ensures a greater safety margin \cite{mathieu_thorium_2006}. While negative
temperature coefficients are attainable with very thermalized spectra,
breeding ratios deteriorated significantly due to parasitic absorption in the
graphite \cite{mathieu_thorium_2006}.
%
\begin{figure}[htb!] 
	\centering
	\includegraphics[width=0.8\textwidth]{MSFR}
	\caption{Schematic view of the MSFR concept. Figure reproduced from
	Brovchenko et al. \cite{brovchenko_design-related_2013}.}
	\label{fig:msfr}
\end{figure}
%
\begin{table}[htb!]
    \small
	\caption{Main specifications of the \gls{MSFR} concept
				\cite{serp_molten_2014}.}
	\centering
	\begin{tabular}{ l r }
		\toprule
		Parameter & Value \\
		\midrule
		Thermal/Electric output [MW$_{\text{th}}$/MW$_{\text{e}}$] & 3000 /
		1500 
		\\
		Salt volume [m$^3$] & 18 \\
		Salt fraction in core & 0.5 \\
		Number of circulation loops & 16 \\
		Nominal flow rate [kg s$^{-1}$] & 18500  \\
		Nominal circulation time [s] & 4.0 \\
		Inlet/outlet temperature [K] & 923 / 1023 \\
		Blanket volume [m$^3$] & 7.3\\
		\bottomrule
	\end{tabular}
	\label{table:msfr}
\end{table}

### 2020-07-04 neutronics.tex
final edits for 2nd draft

This chapter compares key neutronics results between Serpent and Moltres
for a static model of the \gls{MSFR}, i.e. no salt flow, and uniform
temperature distribution to assess the accuracy of the
six-group neutron diffusion model in Moltres on a fast-spectrum reactor. This
exercise also serves as a verification for the group constant data
generated for Moltres from Serpent. For clarity, here are the important
details of the static model.

>>>>

This chapter compares key neutronics results between Serpent and Moltres
for a static model of the \gls{MSFR}, i.e. no salt flow, and uniform
temperature distribution to assess the accuracy of the
six-group neutron diffusion model in Moltres on a fast-spectrum reactor. This
verification exercise builds on a previous study by Lindsay et al.
\cite{lindsay_introduction_2018} that had verified Moltres' neutronics
capabilities with a two-group neutron diffusion model of the \gls{MSRE}.
